### 2017-05-12 2017mergingmerges/section/01-Introduction.tex
very minor type

Secondly, although @sIn1@ occurs twice in the program, at runtime we only want to handle the elements of each stream once. To achieve this, the compiled program must coordinate between the two uses of @sIn1@, so that a new value is read only when both the @group@ and @merge@ operators are ready to receive it. Finally, as the stream length is unbounded, we cannot buffer an arbitrary number of elements read from either stream, or risk running out of local storage space.

>>>>

Secondly, although @sIn1@ occurs twice in the program, at runtime we only want to handle the elements of each stream once. To achieve this, the compiled program must coordinate between the two uses of @sIn1@, so that a new value is read only when both the @group@ and @merge@ operators are ready to receive it. Finally, as the stream length is unbounded, we cannot buffer an arbitrary number of elements read from either stream, or we risk running out of local storage space.

### 2017-05-12 2017mergingmerges/section/02-Processes.tex
very minor type

Figure~\ref{fig:Process:Fused} shows the final result of fusing @group@ and @merge@ together. There are similar rules for handling the other combinations of instructions, but we defer the details to \S\ref{s:Fusion}. The result process has two input streams, @sIn1@ and @sIn2@ and two output streams: @sOut1@ which is @group@'s output stream, and @sOut2@ which is @merge@'s output stream.

To complete the implementation of our example from \S\ref{s:Introduction} we would now proceed to fuse the process from the final line (also a @group@) into this new result process. The order in which pairs of processes are fused together does matter, as does the order in which instructions are interleaved --- we discuss both points further in \S\ref{s:Evaluation}. Although the result process has a single shared heap, the heap bindings from each fused process are guaranteed not to interfere. When we instantiated combinators to create the original source processes we introduced fresh names at that point.

@@

Note that relying on lazy evaluation for an example such as @uniquesUnion@ does not solve the buffering problem. Suppose we converted each of the streams to lazy lists, and used definitions of @group@ and @merge@ that worked over these lists. As @uniquesUnion@ returns a pair of results, there is nothing preventing a consumer from demanding the first list (@sUnique@) in its entirity before demanding any of the elements from the second list (@sUnion@). If this were to happen then the runtime implementation would need to retain all elements of @sIn1@ before demanding any of @sIn2@, causing a space leak. Lazy evaluation is \emph{pull only} meaning that evaluation is driven by the consumer. The space efficiency of our fused program relies critically on the fact that operators can also \emph{push} their result values directly to their consumers, and that the consumers cannot defer the handling of these values.

>>>>

Figure~\ref{fig:Process:Fused} shows the final result of fusing @group@ and @merge@ together. There are similar rules for handling the other combinations of instructions, but we defer the details to \S\ref{s:Fusion}.
The result has two input streams, @sIn1@ and @sIn2@, and two output streams: @sOut1@ from @group@, and @sOut2@ from @merge@. 
The shared input @sIn1@ is pulled by @merge@ at two places, and since both of these need to agree with when @group@ pulls, the @group@ instructions are duplicated at labels @F3@-@F5@ and @F13@-@F15@. The first set of instructions could be simplified to a single @push@ with constant propagation, as @f@ is initially true.

To complete the implementation of our example from \S\ref{s:Introduction} we would now proceed to fuse the process from the final line (also a @group@) into this new result process.
Although the result process has a single shared heap, the heap bindings from each fused process are guaranteed not to interfere: when we instantiated combinators to create the original source processes we introduced fresh names.

% AR: Cut to make room for explanation why two copies of group pull
% The order in which pairs of processes are fused together does matter, as does the order in which instructions are interleaved --- we discuss both points further in \S\ref{s:FusionOrder} and \S\ref{s:Evaluation}.

@@

Note that relying on lazy evaluation for an example such as @uniquesUnion@ does not solve the buffering problem. Suppose we converted each of the streams to lazy lists, and used definitions of @group@ and @merge@ that worked over these lists. As @uniquesUnion@ returns a pair of results, there is nothing preventing a consumer from demanding the first list (@sUnique@) in its entirety before demanding any of the elements from the second list (@sUnion@). If this were to happen then the runtime implementation would need to retain all elements of @sIn1@ before demanding any of @sIn2@, causing a space leak. Lazy evaluation is \emph{pull only} meaning that evaluation is driven by the consumer. The space efficiency of our fused program relies critically on the fact that operators can also \emph{push} their result values directly to their consumers, and that the consumers cannot defer the handling of these values.

### 2017-05-12 2017mergingmerges/section/03-Definition.tex
Appendix: finite stream rules
very minor type

In the grammar of Figure~\ref{fig:Process:Def} the $\InputState$ has three options: @none@, which means no value is currently stored in the associated stream buffer variable, $(@pending@~\Value)$ which gives the current value in the stream buffer variable and indicates that it has not yet been copied into a process-local variable, and @have@ which means the pending value has been copied into a process-local variable. The $\Value$ attached to the @pending@ state is used when specifying the evaluation semantics of processes. When performing the fusion transform the $\Value$ itself will not be known, but we can still reason statically that a process must be in the @pending@ state. When defining the fusion transform in \S\ref{s:Fusion} we will use a version of $\InputState$ with only this staticaly known information.

@@

\item \emph{Feeding} outputs of some processes to the inputs of others. Feeding alternates between Injecting and Advancing. When a process pushes a value to an output channel we attempt to inject this value into all processes that have that same channel as an input. If they all accept it then we then advance their programs as far as they will go, which may cause more values to be pushed to output channels, and so on.
\end{enumerate}

>>>>

In the grammar of Figure~\ref{fig:Process:Def} the $\InputState$ has three options: @none@, which means no value is currently stored in the associated stream buffer variable, $(@pending@~\Value)$ which gives the current value in the stream buffer variable and indicates that it has not yet been copied into a process-local variable, and @have@ which means the pending value has been copied into a process-local variable. The $\Value$ attached to the @pending@ state is used when specifying the evaluation semantics of processes. When performing the fusion transform the $\Value$ itself will not be known, but we can still reason statically that a process must be in the @pending@ state. When defining the fusion transform in \S\ref{s:Fusion} we will use a version of $\InputState$ with only this statically known information.

@@

\item \emph{Feeding} outputs of some processes to the inputs of others. Feeding alternates between Injecting and Advancing. When a process pushes a value to an output channel we attempt to inject this value into all processes that have that same channel as an input. If they all accept it, we then advance their programs as far as they will go, which may cause more values to be pushed to output channels, and so on.
\end{enumerate}

### 2017-05-12 2017mergingmerges/section/05-Evaluation.tex
very minor type

Stream fusion is ultimately performed for practical reasons. We want the fused result program to run faster than unfused source program.

@@

The second file benchmark takes a file and partitions it into two files: one with even-length lines, and one with odd-length lines.
The output lines are also counted.
The three streaming libraries are pull-based, and do not support multiple outputs: the program must be written in a convoluted way or at least partially hand-fused.
Even with hand-fusion, the Pipes and Conduit programs are slower than ours, as well as losing any abstraction benefits from using a streaming library.
The final group in Figure~\ref{fig:bench:all} shows the runtimes for partitioning a 1MB file.

>>>>

Stream fusion is ultimately performed for practical reasons: we want the fused result program to run faster than the original unfused program.

@@

The second file benchmark takes a file and partitions it into two files: one with even-length lines, and one with odd-length lines.
The output lines are also counted.
The three streaming libraries are pull-based, and do not support multiple outputs: the program must be written in a convoluted way or at least partially hand-fused.
Even with partial hand-fusion, the Pipes and Conduit programs are slower than ours, as well as losing any abstraction benefits from using a streaming library.
The final group in Figure~\ref{fig:bench:all} shows the runtimes for partitioning a 1MB file.

### 2017-05-12 2017mergingmerges/section/A1-Finite.tex
very minor type

Figure~\ref{fig:Finite:tryStepPair} shows the modified $\ti{tryStepPair}$ function. This function uses the same heuristics to decide which process to execute when both can progress, but now that the processes can finish with @exit@, we must take care to only finish the fused process once \emph{both} source processes are finished. The (DeferExit1) and (DeferExit2) clauses achieve this by allowing the other process if one is an @exit@. Once both processes are finished, both new clauses will fail while (DeferPull1) succeeds, using the @exit@ from the first process. Another way to think of this is that if either process has work to do, the fused process still has work to do.

@@

Clause (SharedPullInject) applies when both processes need to pull from a shared input. As before we execute a real @pull@, this time with two branches. In the success branch, the input states are set to @pending@ as before. In the closed branch, the input states are set to @closed@ so the next and subsequent pulls take the closed branch.

>>>>

Figure~\ref{fig:Finite:tryStepPair} shows the modified $\ti{tryStepPair}$ function. This function uses the same heuristics to decide which process to execute when both can progress, but now that the processes can finish with @exit@, we must take care to only finish the fused process once \emph{both} source processes are finished. The (DeferExit1) and (DeferExit2) clauses achieve this by forcing the other process to run if one is an @exit@. Once both processes are finished, both new clauses will fail while (DeferPull1) succeeds, using the @exit@ from the first process. Another way to think of this is that if either process has work to do, the fused process still has work to do.

@@

Clause (SharedPullInject) applies when both processes need to pull from a shared input. As before, we execute a real @pull@, this time with two branches. In the success branch, the input states are set to @pending@ as before. In the closed branch, the input states are set to @closed@ so the next and subsequent pulls take the closed branch.
